\documentclass[letterpaper]{report}

\title{Predicting Sea Levels Using the Actuaries Climate Index}
\date{April 7th, 2017}
\author{Jennifer Mince, Karl Maier, Wesley Merrick, Katie Vervack, Toby Duncan}
\begin{document}
	\maketitle

		Our group would like to use the Actuaries Climate Index data set to answer the central question: how do the seasonal components of high temperatures, low temperatures, heavy rain, and drought affect the seasonal sea levels in North American regions and by how much?
		The research we want to conduct is both important and relevant to the societal need of knowing what causes sea levels to rise and predicting future measurements to be better prepared for future problems. Higher sea levels have a significant impact on both local and global environments. Rising levels can cause destructive storms to move closer inland and as a result, there is more coastal flooding. “In the United States, almost 40 percent of the population lives in relatively high-population-density coastal areas, where sea level plays a role in flooding, shoreline erosion, and hazards from storms” (NOAA). Rising seas threaten the infrastructure used by industries and businesses, such as roads, bridges, power plants, and oil wells. The coastline of the United States is heavily populated and the risk of rising sea levels causes many problems.  “Approximately 25 million people live in an area vulnerable to coastal flooding” (EPA). The U.S. economy, transportation, and many ecosystems are threatened by the impending flooding that comes with the rising levels. Integral coastal activities including marine transportation of goods, resource extraction, and tourism help to generate “58% of the national gross domestic product” (EPA). Transportation in the United States is substantially affected by coastal flooding, especially roads. In low-lying communities, the streets are first to be flooded because they are lower than the surrounding land (Titus). The current drainage systems for these roads are not efficient enough to handle increased frequency of flooding due to rising sea levels. 
		Using this data, our group could help this societal need by introducing a model for the rate at which the sea levels are increasing, and calculating factors that contribute to the increase in levels. The data set has an Actuary Climate Index (ACI) created with the components: frequency of temperatures above the 90th percentile (T90), frequency of temperatures below the 10th percentile (T10), maximum rainfall per month in five consecutive days (P), annual maximum consecutive dry days (D), frequency of wind speed above the 90th percentile (W), and sea level changes (S). It is an objective measure of changes in extreme weather and changes in sea level relative to the base period of 1961 through 1990. This index is an educational tool designed to help inform actuaries, public policymakers, and the general public on changes in these measures over recent decades (Actuary Climate Index). Our group will create our own ACI using the seasonal values for T90, T10, P, and D to see how they affect the sea level changes. 
		\newpage
		After researching the various data science studies that were conducted using the ACI data set only one other study was found. The study focused on using the data in the ACI data set to build a risk index for areas where economic losses from  would be higher than other areas. Since there was not much research using the specific ACI data set, I expanded my search to the individual data sets that make up the ACI data set. Even in these further searches, I did not really any data science studies using the data but rather many research papers analyze the data set.
		
		
		
		
	
\end{document}
